Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2016-09-20 paper.tex
list of components to be discussed.
Fix latex

-------------------------------------------------------------------------------------------
\begin{enumerate}
  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{enumerate}

@@

@@

\begin{description}
  \item[Task]     A set of operations to be performed on a computing platform,
                  alongside a description of the properties and dependences of
                  those operations, and indications on how they should be
                  executed and satisfied.
  \item[Workload] A set of tasks that can be executed concurrently, possibly
                  related by a set of relations. For example, tasks of a workload can share one or more input files or communicate during execution.
  \item[Workflow] Set of workloads, related by a set of relations that define
                  the order in which each workload can be executed. Data
                  dependences are the most common relations among workloads,
                  used to define the precedence among their executions. Note
                  that, formally, a workload can have a single task.
\end{description}

@@

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
\input{WMSLiterature}
--------------------------------------------------------------------------------------------------
\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
{enumerate}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{enumerate}

@@

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA developers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:architecture}}
\end{center}
\end{figure}

-------------------------------------------------------------------------------------
\begin{enumerate}
  \item PanDA WMS
  \begin{enumerate}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{enumerate}
  \item PANDA Pilot
\end{enumerate}

>>>>

--------------------------------------------------------------------
\begin{enumerate}
  \item Problem that has been solved: executing many applications on heterogeneous distributed resources.
  \item Why are WMS needed?
  \item Increasing importance of WMS w.r.t complex and scalable applications on heterogeneous distributed resources.
  \item \ldots
  \item Overview/Summary of paper.
\end{enumerate}

@@

--------------------------------------------------------------------

@@

\begin{description}
  \item[Task.] A set of operations to be performed on a computing platform,
  alongside a description of the properties and dependences of those
  operations, and indications on how they should be executed and satisfied.
  \item[Workload.] A set of tasks that can be executed concurrently, possibly
  related by a set of relations. For example, tasks of a workload can share
  one or more input files or communicate during execution.
  \item[Workflow.] Set of workloads, related by a set of relations that define
  the order in which each workload can be executed. Data dependences are the
  most common relations among workloads, used to define the precedence among
  their executions. Note that, formally, a workload can have a single task.
\end{description}

@@

% ----------------------------------------------------------------------------
% III
% ----------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
\input{WMSLiterature}
--------------------------------------------------------------------
\begin{enumerate}
  \item Other workload management systems (WMS) (not exclusive)
  \begin{enumerate}
    \item glidein-WMS
    \item Dirac
    \item ALICE-AlieN
  \end{enumerate}
{enumerate}
  \item Comparison and contrast
  \item Core features of a general WMS, i.e., minimal complete model of a WMS.
\end{enumerate}

@@

\subsection{Use of Supercomputers with PanDA}
The idea is to extend the ATLAS Computing Model beyond the Grid into the domains of supercomputers. Modern High Performance Computing  (HPC) involves  a very large number of worker nodes  connected  through a  high-speed   network. The worker nodes can have  multicore CPUs that can be augmented with massively parallel Graphics Processing Units (GPUs) or other types of coprocessors. On an HPC machine, a  typical job is highly parallel and connected,  where each core calculates  a small part of the problem and coordinates the activity via message passing interface (MPI). While this is different from the HEP computing paradigm, where jobs are independent, it still shares common  features  such as the use of parallelization. It is not a requirement  that HPC machines are able to run any possible task, nor is it relevant how many kinds of job types that can be run. What matters is the total number of cycles that can be offloaded from the Grid. Standard  ATLAS  workflow  can not be easily ported to supercomputers   due  to   several   complications such  as specialized worker  node  setups, no  outbound network connections,  limited memory per node, custom operating systems,  etc. A reorganization  of the standard  workflow is therefore needed.
The most suitable task for HPCs in HEP is event generation. Event generators are mostly computational,  stand-alone code, with few input requirements.  There is no need  to stage-in much data. The event generation  in ATLAS corresponds  to
10-15\% of all jobs on the Grid. A more difficult workload to adapt  for HPCs is detector  simulation. Simulation tasks tend to be tied to the framework of a particular simulation and often require  database access for geometry and relevant run conditions.  However,  there is a  lot to be gained  since simulation  tasks correspond to 2/3 of the Grid capacity.
\subsection{Interfacing PanDA with Titan}
The Titan supercomputer, current number two (number one from November 2012 until June 2013) on the Top 500 list is located at the Oak Ridge Leadership Computing  Facility (OLCF) in Oak Ridge National Laboratory, US. Titan, was the first large-scale system to use  a hybrid architecture that utilizes worker nodes with both AMD 16-core Opteron 6274 CPUs and NVIDIA Tesla K20 GPU accelerators.
Integration with Titan is the current focus for PanDA developers. The project aims to integrate Titan with the PanDA system using an updated PanDA Pilot that runs on the front-end node and submits ATLAS payloads to the worker nodes using the local batch system (PBS) via the SAGA (Simple API for Grid Applications) interface \cite{SAGA}. This solves several complications of  running on HPC worker nodes,  including the lack of connectivity to outside world. The pilot can communicate with the PanDA server from the front-end machine. The interactive front-end  machines and the worker nodes  use  a shared  file system  which makes  it  easy  for the pilot to stage-in  any input files that are required by the payload and stage-out the produced output files after the payload has finished at the end of the job. ATLAS Tier 1 computing  center at Brookhaven National Lab is currently  used for data stage out from Titan but in principle  that can be any Grid site.
1) HPC Backfill: HPC facilities are geared towards  large scale jobs by design. Time allocation on an HPC is competitive and large projects are often preferred. About 10% of capacity on a  typical HPC machine  is unused  due to mismatches between job sizes and available  resources. The worker nodes
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/PandaInterfaceWithHPC.png}
\caption{Schematic view of the PanDA system\label{fig:interface}}
\end{center}
\end{figure}

--------------------------------------------------------------------
\begin{enumerate}
  \item PanDA WMS
  \begin{enumerate}
    \item State Diagrams
    \item Control Flow Diagrams
    \item Software Module and Architecture diagrams.
  \end{enumerate}
  \item PANDA Pilot
\end{enumerate}
